---
title: "Test"
author: "Dana Aspel"
date: "2025-08-19"
output: html_document
runtime: shiny
---

This app uses:

-   **`faithful`** --- a built-in dataset in base R (no package needed)
-   **`shinyAppDir()`** --- a function from the shiny package
-   **`system.file()`** --- a base R function

Packages and dependencies:

-   **`base R`** (for everything else)
-   **`shiny`**
-   **`dplyr`**
-   **`tidyr`**
-   **`data.table`**
-   **`lattice`**
-   **`forecast`**
-   

```{r setup, include=FALSE}
library(shiny)
library(dplyr) 
library(tidyr) 
library(data.table)
library(lattice)
# library(forecast)
```

This R Markdown document is made interactive using Shiny. This uses the dataset Faithful, which contains a record of Old Faithful geyser, Yellowstone, eruptions.

## Data Prep

The dataset is first filtered to include only geyser eruptions longer than three minutes. These durations are then rounded to the nearest half-minute to simplify grouping. The number of occurrences for each rounded duration is counted, and the results are reshaped into a wide format for easier comparison.

```{r data_summary, echo=FALSE}

output$summaryTable <- renderTable({ faithful %>% filter(eruptions > 3) %>%
mutate(rounded = round(eruptions * 2) / 2) %>% count(rounded) %>% pivot_wider(names_from = rounded, values_from = n, values_fill = 0) %>% as.data.table() })
tableOutput("summaryTable")
```

## Eruptions

Old Faithful geyser eruption duration.

```{r eruptions, echo=FALSE}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

## Statistical Application

This uses the system file 06_tabsets to create a Shiny application.

```{r tabsets, echo=FALSE}
shinyAppDir(
  system.file("examples/06_tabsets", package = "shiny"),
  options = list(
    width = "100%", height = 550
  )
)
```

## Eruption vs. Waiting Time (Lattice)

This plot uses the \`lattice\` package to show the relationship between eruption duration and waiting time between eruptions.

```{r lattice_plot, echo=FALSE}
output$latticePlot <- renderPlot({
  xyplot(waiting ~ eruptions, data = faithful,
         xlab = "Eruption Duration (minutes)",
         ylab = "Waiting Time (minutes)",
         main = "Old Faithful Eruption vs. Waiting Time")
})

plotOutput("latticePlot")
```

## Forecasting Section

## Failed dependencies

| Package  | Sub-Dependency | Error Code    | Req. Version |
|----------|----------------|---------------|--------------|
| plotly   | MASS           | exit status 1 | R \>= 4.4.0  |
| forecast | MASS           | exit status 1 | R \>= 4.4.0  |
| ggplot2  | MASS           | exit status 1 | R \>= 4.4.0  |
| car      | MASS           | exit status 1 | R \>= 4.4.0  |
| caret    | MASS           | exit status 1 | R \>= 4.4.0  |
