---
title: "Test-1"
author: "Dana Aspel"
date: "2025-08-19"
output: html_document
runtime: shiny
---

This app uses:

-   **`faithful`** --- a built-in dataset in base R (no package needed)
-   **`shinyAppDir()`** --- a function from the shiny package
-   **`system.file()`** --- a base R function

Packages and dependcies:

-   **`shiny`** (explicitly used)

-   **`base R`** (for everything else)

```{r setup, include=FALSE}
library(shiny)
```

This R Markdown document is made interactive using Shiny. This uses the dataset Faithful, which contains a record of Old Faithful geyser, Yellowstone, eruptions.

## Eruptions

Old Faithful geyser eruption durations.

```{r eruptions, echo=FALSE}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

## Statistical Application

This uses the system file 06_tabsets to create a Shiny application.

```{r tabsets, echo=FALSE}
shinyAppDir(
  system.file("examples/06_tabsets", package = "shiny"),
  options = list(
    width = "100%", height = 550
  )
)
```

In all of R code chunks above the `echo = FALSE` attribute is used. This is to prevent the R code within the chunk from rendering in the document alongside the Shiny components.

## Renv dependencies

| Source                                                           | Package   | Require Version | Dev   |
|-----------------------------|---------------|---------------|---------------|
| C:/Users/aspeld/Repos/Posit-Test/Posit-Connect-Test-R/Test-1.Rmd | rmarkdown |                 | FALSE |
| C:/Users/aspeld/Repos/Posit-Test/Posit-Connect-Test-R/Test-1.Rmd | shiny     |                 | FALSE |
| C:/Users/aspeld/Repos/Posit-Test/Posit-Connect-Test-R/Test-1.Rmd | shiny     |                 | FALSE |
